An Empirical Analysis of the Role of Energy in Economic Growth

### 2013-04-13 EconGrowth-R-Functions.Rnw
Removed many debugging print statements in the functions.

print("Inside cobbDouglasData")
print(paste("energyType =", energyType))

@@

print("After if ! haveDataCD")

@@

print("After cobbDouglasModel")

@@

print("Inside if is.na(energyType)")

@@

###########################################
# Creates a number of rows in a data.frame that contain information 
# about the coefficients of a Cobb-Douglas model for countryAbbrev.
# Each parameter has its own row with confidence intervals.
# The country name is in a column. Which parameter is involved is
# also in a column.
# 
# The return type is a data.frame.
##
cobbDouglasCountryRowsForParamsGraph <- function(countryAbbrev, energyType){
print("Inside cobbDouglasCountryRowsForParamsGraph")
  #Create three rows, one for each parameter. Each row is a data.frame so that it is plottable!
  if (is.na(energyType)){
print("Inside is.na(energyType)")
    valueRow <- "CD"
    dataCD <- cobbDouglasData(countryAbbrev=countryAbbrev, energyType=NA)
print("After cobbDouglasData")
  } else {
    valueRow <- "CDe"
    dataCD <- cobbDouglasData(countryAbbrev=countryAbbrev, energyType=energyType)
  }
  # These rows are common to the "with" and "without" energy cases.
  lambdaRow <- data.frame(country = countryAbbrev, 
                          parameter = "lambda", 
                          lowerCI = dataCD["-95% CI", "lambda"], 
                          value = dataCD[valueRow, "lambda"], 
                          upperCI = dataCD["+95% CI", "lambda"])
  alphaRow <- data.frame(country = countryAbbrev, 
                         parameter = "alpha",
                         lowerCI = dataCD["-95% CI", "alpha"],
                         value = dataCD[valueRow, "alpha"],  
                         upperCI = dataCD["+95% CI", "alpha"])
  betaRow <- data.frame(country = countryAbbrev, 
                        parameter = "beta",
                        lowerCI = dataCD["-95% CI", "beta"],
                        value = dataCD[valueRow, "beta"], 
                        upperCI = dataCD["+95% CI", "beta"])
  if (is.na(energyType)){
    table <- rbind(lambdaRow, alphaRow, betaRow)
  } else {
    gammaRow <- data.frame(country = countryAbbrev, 
                           parameter = "gamma",
                           lowerCI = dataCD["-95% CI", "gamma"],
                           value = dataCD[valueRow, "gamma"], 
                           upperCI = dataCD["+95% CI", "gamma"])
    table <- rbind(lambdaRow, alphaRow, betaRow, gammaRow)
  }
  return(table)
}

@@

print("Inside createCDParamsGraph")

@@

print("After lapply(cobbDouglasCountryRowsForParamsGraph")

@@

print("After cobbDouglasPredictionsColumn(energyType=NA)")

@@

print("Before xyplot")

@@

print("After xyplot")

@@

######################################
# Loads data and trims rows where necessary.
# Also, replaces the name of the desired energy column with "iEToFit".
# If you set energyType=NA, "iEToFit" will not appear in the column names, because
# it is assumed that you want a CES fit without an energy term.
# Returns a data.frame with the loaded data.
##
loadAndPrepDataForCES <- function(countryAbbrev, energyType){
print(paste("Inside loadAndPrepDataForCES. energyType =", energyType))
  dataTable <- loadData(countryAbbrev)
  # Trim data from China and South Africa that is missing labor information for 2011.
  dataTable <- subset(dataTable, !is.na(iLabor))
  if (is.na(energyType)){
print("Inside is.na(energyType) in loadAndPrepDataForCES")
    # We're done. Don't have to deal with special issues with energyType.
    return(dataTable)
  }
print("After is.na(energyType) block")
  # We have an energyType argument. Do some additional checking.
  if (energyType == "U"){
    # Trim the dataset to include only those years for which U is available.
    dataTable <- subset(dataTable, !is.na(iU))
  }
  # We need to do the CES fit with the desired energyType.
  # To achieve the correct fit, we'll change the name of the desired column
  # to "iEToFit" and use "iEToFit" in the nls function.
  dataTable <- replaceColName(dataTable, energyType, "iEToFit")
  return(dataTable)
}

@@

print(paste("Inside cesModel. energyType =", energyType))

@@

print("Inside is.na(energyType)")

@@

#########################
# Takes the CES fitted models and creates per-country predictions for them.
# Returns a data.frame with the predictions.
# If energyType=NA, the CES model without energy will be used.
##
cesPredictions <- function(countryAbbrev, energyType){
print(paste("Inside cesPredictions. energyType =", energyType))
  # Can't make predictions for any of CN, ZA, SA, IR, TZ, or ZM if we're interested in U
  if (!(haveDataCES(countryAbbrev, energyType))){
    # If we don't have data for this combination of countryAbbrev and energyType, 
    # return a column of NAs if the above conditions have been met.
    nRows <- 21 # All of these countries need 21 rows.
    df <- as.data.frame(matrix(NA, ncol = 1, nrow = nRows))
    colnames(df) <- "pred"
    return(df)
  }
  model <- cesModel(countryAbbrev=countryAbbrev, energyType=energyType)
  pred <- fitted(model)
  df <- data.frame(pred)
  # Pad with rows as necessary
  df <- padRows(countryAbbrev, df)
  return(df)
}

#########################
# Takes the CES fitted models and creates a single column of predicted GDP values
# that corresponds, row for row, with the AllData.txt file.
# If energyType=NA is specified, the CES model without energy will be used for the predictions.
##
cesPredictionsColumn <- function(energyType){
print(paste("Inside cesPredictionsColumn. energyType =", energyType))
  out <- do.call("rbind", lapply(countryAbbrevs, cesPredictions, energyType=energyType))  
print("After do.call(rbind, etc.)")
  if (is.na(energyType)){
    colnames(out) <- "predGDP"
  } else {
    colnames(out) <- c(paste("predGDP", energyType, sep=""))
  }
  return(out)
}

@@

print("Inside createCESLatticeGraph.")

@@

print("Made it to after collecting all the predictions.")

@@

print("Inside cesData")
print(paste("is.na(energyType) =", is.na(energyType)))

@@

print(paste("wantEnergyTerm =", wantEnergyTerm))

>>>>

@@

@@

@@

@@

###########################################
# Creates a number of rows in a data.frame that contain information 
# about the coefficients of a Cobb-Douglas model for countryAbbrev.
# Each parameter has its own row with confidence intervals.
# The country name is in a column. Which parameter is involved is
# also in a column.
# 
# The return type is a data.frame.
##
cobbDouglasCountryRowsForParamsGraph <- function(countryAbbrev, energyType){
  #Create three rows, one for each parameter. Each row is a data.frame so that it is plottable!
  if (is.na(energyType)){
    valueRow <- "CD"
    dataCD <- cobbDouglasData(countryAbbrev=countryAbbrev, energyType=NA)
  } else {
    valueRow <- "CDe"
    dataCD <- cobbDouglasData(countryAbbrev=countryAbbrev, energyType=energyType)
  }
  # These rows are common to the "with" and "without" energy cases.
  lambdaRow <- data.frame(country = countryAbbrev, 
                          parameter = "lambda", 
                          lowerCI = dataCD["-95% CI", "lambda"], 
                          value = dataCD[valueRow, "lambda"], 
                          upperCI = dataCD["+95% CI", "lambda"])
  alphaRow <- data.frame(country = countryAbbrev, 
                         parameter = "alpha",
                         lowerCI = dataCD["-95% CI", "alpha"],
                         value = dataCD[valueRow, "alpha"],  
                         upperCI = dataCD["+95% CI", "alpha"])
  betaRow <- data.frame(country = countryAbbrev, 
                        parameter = "beta",
                        lowerCI = dataCD["-95% CI", "beta"],
                        value = dataCD[valueRow, "beta"], 
                        upperCI = dataCD["+95% CI", "beta"])
  if (is.na(energyType)){
    table <- rbind(lambdaRow, alphaRow, betaRow)
  } else {
    gammaRow <- data.frame(country = countryAbbrev, 
                           parameter = "gamma",
                           lowerCI = dataCD["-95% CI", "gamma"],
                           value = dataCD[valueRow, "gamma"], 
                           upperCI = dataCD["+95% CI", "gamma"])
    table <- rbind(lambdaRow, alphaRow, betaRow, gammaRow)
  }
  return(table)
}

@@

@@

@@

@@

@@

@@

######################################
# Loads data and trims rows where necessary.
# Also, replaces the name of the desired energy column with "iEToFit".
# If you set energyType=NA, "iEToFit" will not appear in the column names, because
# it is assumed that you want a CES fit without an energy term.
# Returns a data.frame with the loaded data.
##
loadAndPrepDataForCES <- function(countryAbbrev, energyType){
  dataTable <- loadData(countryAbbrev)
  # Trim data from China and South Africa that is missing labor information for 2011.
  dataTable <- subset(dataTable, !is.na(iLabor))
  if (is.na(energyType)){
    # We're done. Don't have to deal with special issues with energyType.
    return(dataTable)
  }
  # We have an energyType argument. Do some additional checking.
  if (energyType == "U"){
    # Trim the dataset to include only those years for which U is available.
    dataTable <- subset(dataTable, !is.na(iU))
  }
  # We need to do the CES fit with the desired energyType.
  # To achieve the correct fit, we'll change the name of the desired column
  # to "iEToFit" and use "iEToFit" in the nls function.
  dataTable <- replaceColName(dataTable, energyType, "iEToFit")
  return(dataTable)
}

@@

@@

@@

#########################
# Takes the CES fitted models and creates per-country predictions for them.
# Returns a data.frame with the predictions.
# If energyType=NA, the CES model without energy will be used.
##
cesPredictions <- function(countryAbbrev, energyType){
  # Can't make predictions for any of CN, ZA, SA, IR, TZ, or ZM if we're interested in U
  if (!(haveDataCES(countryAbbrev, energyType))){
    # If we don't have data for this combination of countryAbbrev and energyType, 
    # return a column of NAs if the above conditions have been met.
    nRows <- 21 # All of these countries need 21 rows.
    df <- as.data.frame(matrix(NA, ncol = 1, nrow = nRows))
    colnames(df) <- "pred"
    return(df)
  }
  model <- cesModel(countryAbbrev=countryAbbrev, energyType=energyType)
  pred <- fitted(model)
  df <- data.frame(pred)
  # Pad with rows as necessary
  df <- padRows(countryAbbrev, df)
  return(df)
}

#########################
# Takes the CES fitted models and creates a single column of predicted GDP values
# that corresponds, row for row, with the AllData.txt file.
# If energyType=NA is specified, the CES model without energy will be used for the predictions.
##
cesPredictionsColumn <- function(energyType){
  out <- do.call("rbind", lapply(countryAbbrevs, cesPredictions, energyType=energyType))  
  if (is.na(energyType)){
    colnames(out) <- "predGDP"
  } else {
    colnames(out) <- c(paste("predGDP", energyType, sep=""))
  }
  return(out)
}

@@

@@

@@

@@

### 2013-04-13 EconGrowth.Rnw
Removed many debugging print statements in the functions.

\item{Heun: Convert to using NA everywhere for models without energy}

@@

@@

\item \st{Heun: Re-organize CES tables into (a) $\lambda$, $\delta$, $\sigma$ and (b) $\gamma$, $\delta_1$, $\sigma_1$}

>>>>

\item{Heun: Convert the presentation to use energyType=NA for models without energy}

@@

\item{\st{Heun: Convert to using NA everywhere for models without energy}}

@@

\item{\st{Heun: Re-organize CES tables into (a) $\lambda$, $\delta$, $\sigma$ and (b) $\gamma$, $\delta_1$, $\sigma_1$}}
